---
title: "Inferential Analysis of Exponential Distribution through Simulation by CLT"
author: "Gowri Shankar"
date: "09/09/2015"
output: html_document
hitheme: tomorrow
mode: selfcontained
highlighter: highlight.js
framework: io2012
url:
  assets: ../../assets
  lib: ../../librariesNew
widgets: mathjax
---

## Overview

Exponential distribution is a process in which events occur continuously and independently
at a constant average rate.

 - Probability Density Function: **E[X] = λ e^(−λx)**
 - Cumulative Distribution Function: **Var[X] = 1 − e^(−λx)**
 - Mean: **1/λ**
 - Variance: **1/λ^2** i.e Standard Deviation: **1/λ**

This is an inferential report on the simulation analysis conducted on exponential distribution
data using central limit theorem. It proves theoretical mean and variance quoted above with simulated data.

## Simulations
Exponentially distributed data is generated using R's **rexp** function. 3 kinds of simulations are done to prove theoretical mean **1/λ** and simulated mean converges.

 - for 1000 iteration with **rate** parameter of **0.2** cumulative mean was calculated
 - Simulation for 1000 iterations with 40 exponentials
 - Law of Large Numbers test with 400 exponentials
 - a histogram is drawn to show the density of distributed data is normal by portraying a bell curve


```{r}
nosims <- 1000; 
lambda <- 0.2
n = 40
set.seed(123456)
```

## Sample Mean vs Theoretical Mean(1000 Iterations)


```{r, fig.height=5, fig.width=10}
means <- cumsum(rexp(nosims, lambda)) / (1  : nosims); library(ggplot2)
g <- ggplot(data.frame(x = 1 : nosims, y = means), aes(x = x, y = y)) 
g <- g + geom_hline(yintercept = 1/lambda,  color = I('blue'), size = 1.5) + geom_line(size = 2) 
g <- g + labs(x = "Number of iterations", y = "Cumulative mean")
g
```

Above plot demonstrates the cumulative mean of sample mean converges eventually at theoretical mean **1/λ = `r 1/lambda`**


## Sample Mean vs Theoretical Mean(1000 Iterations of 40 exponentials)

```{r, fig.height=5, fig.width=10}

sim_data = data.frame(x = sapply(1:nosims, function(x) { rexp(n, lambda) }))
means = data.frame(x = sapply(sim_data[, 1:nosims], mean))
qplot(x, y, data=data.frame(x = 1:nosims, y = means$x)) + 
    geom_smooth() + 
    geom_hline(yintercept = 1/lambda) + 
    labs(x = "Number of obs", y = "Cumulative mean")
```

## Law of Large Numbers(1000 Iterations of 400 exponentials)
No let us look at what happens when number of exponentials are 400 and iterated for 1000 times.

```{r, fig.height=5, fig.width=10}
means = data.frame(x = sapply(1:nosims, function(x) {
    mean(rexp(400, lambda))
}))
qplot(x, y, data=data.frame(x = 1:nosims, y = means$x)) + 
    geom_smooth() + 
    geom_hline(yintercept = 1/lambda) + 
    labs(x = "Number of obs", y = "Cumulative mean")
```

Above plot once again proves the sample mean converges at theoretical mean. Now the error is very less, distribution quite closer to **1/λ = `r 1/lambda`**

## Sample Variance vs Theoretical Variance
Let us take Standard Deviation instead of Variance, so that our plots are uniform.

```{r, fig.height=5, fig.width=10}

sds = sapply(sim_data[, 1:nosims], sd)
qplot(x, y, data=data.frame(x = 1:nosims, y = sds)) + geom_smooth() + geom_hline(yintercept = 1/lambda) + labs(x = "Number of iteration", y = "Standard Deviation")
```

This plot also shows the data is dense around **1/λ = `r 1/lambda`**

## Distribution

```{r, fig.height=5, fig.width=10}
ggplot(data = means, aes(x = x)) + 
    geom_histogram(aes(y=..density..), fill = I('#aaccff'), color = I('blue')) +
    stat_function(fun = dnorm, arg = list(mean = 1/lambda, sd = sd(means$x)), color = I('green'), size = 1.5) + 
    geom_vline(xintercept = 1/lambda, size = 1, color = I('red')) 

```

Above plot is quite Gaussian with dense distribution around the theorteical mean **1/λ = `r 1/lambda`** in red line with sample standard deviation of **1/λ = `r mean(sds)`** in green line.
Sample mean turned out to be **`r mean(means$x)`**
