```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F, results='hide'}
library(UsingR)
library(manipulate)
```

## Least Square Equation

Consider the data set given below

x <- c(0.18, -1.54, 0.42, 0.95)

and weights given by

w <- c(2, 1, 3, 1)

Give the value of 'mu' that minimizes the least squares equation 

SUM(w*(x - mu)^2)


```{r}
mse = sapply(c(0.0025, 1.077, 0.3, 0.1471), function(mu) {
    x <- c(0.18, -1.54, 0.42, 0.95)
    w <- c(2, 1, 3, 1)
    mse <- sum(w*(x - mu)^2)
    return(mse)
})
mse
```

## Regression through Origin
Consider the following data set

x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)

y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

Fit the regression through the origin and get the slope treating y as the
the outcome and x as the regressor. (Hint, do not center the data since 
we want regression through the origin, not through the means of the data)

```{r}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

# lm(I(x - mean(x)) ~ I(y - mean(y) ))
lm(I(y - mean(y)) ~ I(x - mean(x)))
``` 

### Plot and find beta using manipulate

```{r}
x1 <- x - mean(x)
y1 <- y - mean(y)
freqData <- as.data.frame(table(x1, y1))
names(freqData) <- c("x", "y", "freq")
freqData$x <- as.numeric(as.character(freqData$x))
freqData$y <- as.numeric(as.character(freqData$y))
myPlot <- function(beta){
    g <- ggplot(filter(freqData, freq > 0), aes(x = y, y = x))
    g <- g  + scale_size(range = c(2, 20), guide = "none" )
    g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
    g <- g + geom_point(aes(colour=freq, size = freq))
    g <- g + scale_colour_gradient(low = "lightblue", high="white")                     
    g <- g + geom_abline(intercept = 0, slope = beta, size = 3)
    mse <- mean( (y1 - beta * x1) ^2 )
    g <- g + ggtitle(paste("beta = ", beta, "mse = ", round(mse, 3)))
    g
}
# manipulate(myPlot(beta), beta = slider(-2, 0, step = 0.1))
```

### Verify the result with options
```{r}
mse1 = sapply(c(-0.04462, 0.8263, -1.713, 0.59915), function(beta) {
    mse <- mean((y1 - x1*beta)^2)
    return(mse)
})
mse1
```

## Predict through Fitting a Regression Model

Do data(mtcars) from the datasets package and fit the regression model with mpg as the outcome and weight as the predictor. Give the slope coefficient.

```{r}
head(mtcars, 3)
lm(I(mpg - mean(mpg)) ~ I(wt - mean(wt)) - 1, mtcars)
```
