---
title: "Activity Monitoring"
author: "Gowri Shankar"
date: "07/22/2015"
output: html_document
---

Activity monitoring devices like Fitbit, Jawbone Up, Nike Fuelband are quite popular among health conscious individuals and tech geeks. These devices provides thier user a quantified output their movements and activities. Use cases range from Sleep Monitoring to No. of steps a person walked on a given period of time.

This analysis is done from the data collected from one such device. The device had collected data at 5 minutes interval throughout the day over a period of 2 months. 

## Synopsis
The report consummates 8 different operations and answers few key questions



```{r include = FALSE}
# Load required libraries
library(data.table)
library(dplyr)
library(lubridate)
library(plyr)
library(reshape2)
library(ggplot2)
library(ggExtra)
library(gtable)
```

## Loading and preprocessing the data
Data is read from a csv file and transformed into table data frame.
Using 'mutate' function date in string format is converted toe Time(POSIXct) format

```{r}
data <- read.table("activity.csv", header=T, sep=',', na.strings="NA", 
                   stringsAsFactors = FALSE)
# Convert the data to table data and remove from the memory
new_data <- tbl_df(data)
rm(data)


# Mutate so that the format of Date and Time(POSIXct) column is no longer a string
new_data <- mutate(new_data, date = as.Date(date, format="%Y-%m-%d"))
```    

## What is mean total number of steps taken per day?
Mean and Median total number of steps taken per day is calculated using 'ddply' function
```{r}
calcStepsPerDay <- function(data) {
    stepsPerDay <- ddply(
        data, 
        .(date), 
        function(x) {  
            sum(x$steps, na.rm = T)
        }
    )
    return(stepsPerDay)
}
stepsPerDay <- calcStepsPerDay(new_data)
plot(x = stepsPerDay$date, y = stepsPerDay$V1, type= "h", main = "Total No of Steps Taken Each Day",
     xlab = "Days", ylab = "Steps(total)")
mean <- mean(stepsPerDay$V1)
median <- median(stepsPerDay$V1)
```
Mean and Median are `r mean` and `r median` respectively

## What is the average daily activity pattern?
A time series plot is made across 5 minute interval throughout the day and the average number of steps taken.
```{r}
average <- ddply(
    new_data, 
    .(interval), 
    function(x) {  
        mean(x$steps, na.rm = T)
    }
)
plot(x = average$interval, y = average$V1, type = "l", main = "Average No. Of Steps Taken(Time Series)",
     xlab = "5 Mins Interval", ylab = "Steps(Average)")
maxInterval <- average[which.max(average$V1),]
```
The average steps taken is calculated across 2 months of days for that particular 5 minute interval

It has been noticed that the `r maxInterval$interval`th 5 minute interval contains the maximum number steps across all days.


## Imputing missing values
While reading the data from data source, cells with no values are replaced with 'NA' this section establishes a strategy for filling the missing data.

### Impute Strategy
Missing variables are replaced with the population mean of the given interval.

#### Steps
* Identify the rows having NAs using complete cases.
* Reuse the average data calculated for every 5 minute interval
* Iterate through all the rows and find the NA rows
* Replace NAs with calculated averages.

```{r}
na_rows <- complete.cases(new_data)
noOfMissingRows <- 0
for(i in 1 : nrow(new_data)) {
    if(na_rows[i] == FALSE) {
        index <- match(new_data[i, "interval"], average$interval)
        new_data[i, "steps"] <- average[index, "V1"]
        noOfMissingRows <- noOfMissingRows + 1
    } 
}
stepsPerDay <- calcStepsPerDay(new_data)
plot(x = stepsPerDay$date, y = stepsPerDay$V1, type= "h", main = "Total No of Steps Taken Each Day After Imputing", xlab = "Days", ylab = "Steps(total)")
mean <- mean(stepsPerDay$V1)
median <- median(stepsPerDay$V1)
```
In `r noOfMissingRows` rows data is not recorded.

New Mean and Median are `r mean` and `r median` respectively. These values are quite different from the actual values. Impact of imputing resulted in same median and mean.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
days <- c("Saturday", "Sunday")
weekdays <- filter(new_data, is.na(match(weekdays(date), days)))
weekends <- filter(new_data, !is.na(match(weekdays(date), days)))

aveWeekdayInterval <- ddply(
    weekdays, 
    .(interval), 
    function(x) {  
        mean(x$steps, na.rm = T)
    }
)

aveWeekendInterval <- ddply(
    weekends, 
    .(interval), 
    function(x) {  
        mean(x$steps, na.rm = T)
    }
)
p1 <- ggplot(aveWeekdayInterval, aes(interval, V1)) + geom_line() + ggtitle("Average number of steps taken     across weekdays and weekends") + xlab("Interval(Weekdays)") + ylab("Average Steps")
p2 <- ggplot(aveWeekendInterval, aes(interval, V1)) + xlab ("Interval(Weekends)") + ylab("Average Steps") + geom_line()

gl = lapply(list(p1,p2), ggplotGrob)     
g = do.call(rbind, c(gl, size="first"))
g$widths = do.call(unit.pmax, lapply(gl, "[[", "widths"))

grid.newpage()
grid.draw(g) 
```

The above two plots compares the weekdays vs weekend activities.
