---
title: "Sytematic Analysis on Storm Data (1950 - 2011)"
author: "Gowri Shankar"
date: "07/21/2015"
output: html_document
---

This report consummates the analysis and findings of storm data data collected between the years 1950 and 2011. Data is collected by National Weather Service and National Climatic Data Center which explores the severe weather events which can result in public health and economic problems for communities and municipalities.

### Synopsis
This report consist of 2 major sections which has

* A summary of how the data is processed

* Inferences and the Results arrived 

    * Net Fatalities due to various events
    * Net Injuries due to various events
    * Economic loss because of Crop Damage
    * Economic loss due to Property Damage
```{r}
library(data.table)
library(dplyr)
library(lubridate)
library(plyr)
library(reshape2)
library(ggplot2)
```
### Data Processing
* Data is downloaded from the repository and unzipped to have plain csv file.
* Plain CSV file is loaded using read.csv method.
* While loading the CSV data, strings are interpreted as factors
* Once loaded, data is converted to data frame format(tbl_df)
* To avoid multiple load, presence of data is varified using "exists" method

#### Data Tranformation
Data is tranformed from flat csv loaded one to Data Frame format for quick analysis using ddply functions

```{r}
library(data.table)
library(dplyr)
library(lubridate)
library(plyr)
library(reshape2)
library(ggplot2)

if (!(exists("storm_data"))) {
    print("Storm Data Loading...")
    if (!(exists("csv_data"))) {
        csv_data <- read.csv("StormData.csv", sep = ",", stringsAsFactors = F)
    }
    storm_data <- tbl_df(csv_data)
    rm(csv_data)
}
```


### Results
This section provides an exhaustive summary of the fatalities, injuries and losses incurred due to natural calamities listed in the source data.

#### Events Most Harmful to Population Health
This section portrays the top 5 event which causes human death and injuries respectivelys.

**Fatalities**
Tornados resulted in more than 5500 deaths and the next in the list is Excessive Heat which is quite less than half of the death caused by tornados.
```{r}
calcNetFatalitiesByEvent <- function(data) {
    fatalitiesByEvent <- ddply(
        storm_data, 
        .(EVTYPE), 
        function(x) {  
            sum(x$FATALITIES, na.rm = T)
        }
    )
    return(fatalitiesByEvent[order(-fatalitiesByEvent$V1), ])
}
netFatalities <- calcNetFatalitiesByEvent(storm_data)

tmpData1 <- netFatalities[1:5, ]

p <- qplot(
    EVTYPE, V1, stat="identity", data = tmpData1, geom = "bar",
    xlab = "Event Type",
    ylab = "No of Deaths",
    main = "Net Fatalities of Top 5 Events",
    fill = factor(EVTYPE)
) 
print(p)

print(tmpData1)
```

**Injuries**
Injuries caused by tornados are way high compared to any other natural calamity. 
```{r}
calcNetInjuriesByEvent <- function(data) {
    injuriesByEvent <- ddply(
        storm_data, 
        .(EVTYPE), 
        function(x) {  
            sum(x$INJURIES, na.rm = T)
        }
    )
    return(injuriesByEvent[order(-injuriesByEvent$V1), ])
}

netInjuries <- calcNetInjuriesByEvent(storm_data)

tmpData2 <- netInjuries[1:5, ]

q <- qplot(EVTYPE, V1, stat="identity", data = tmpData2, geom = "bar",
    xlab = "Event Type",
    ylab = "No of Injuries",
    main = "Net Injuries of Top 5 Events",
    fill = factor(EVTYPE)
)
print(q)

print(tmpData2)
```

#### Events Causing Maximum Economic Loss
This section portrays the top 10 events which causes maximum economic losses.

**Crop Damage**
Hail storms and flash floods result in high crop damage. The losses due to hail storm is more than 3 times its next damage causing event which is flash floods.

```{r}
calcNetCropDamage <- function(data) {
    cropDamageByEvent <- ddply(
        storm_data, 
        .(EVTYPE), 
        function(x) {  
            sum(x$CROPDMG, na.rm = T)
        }
    )
    return(cropDamageByEvent[order(-cropDamageByEvent$V1), ])
}

cropDamage <- calcNetCropDamage(storm_data)
print(head(cropDamage, 10))
```

**Property Damage**
Tornados has resulted in highest property damage of order 3.2 billion.

```{r}
calcNetPropertyDamage <- function(data) {
    propertyDamageByEvent <- ddply(
        storm_data, 
        .(EVTYPE), 
        function(x) {  
            sum(x$PROPDMG, na.rm = T)
        }
    )
    return(propertyDamageByEvent[order(-propertyDamageByEvent$V1), ])
}

propertyDamage <- calcNetPropertyDamage(storm_data)
print(head(propertyDamage, 10))
```
