What is the variance of the distribution of the average an IID draw of n observations from a population with mean μ and variance σ2.

sigma.square/n

## Q2

Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?

```{r}
pnorm(70, mean = 80, sd = 10, lower.tail = TRUE)
```


## Q3: Done

Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?

Mean: 1100
SD: 75
95th Percentile = ?
```{r}

qnorm(0.95, mean = 1100, sd = 75)

```


## Q4

Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean?

Mean: 1100
SD: 75
n = 100
95th Percentile = ?

```{r}

new_mu <- mean(rnorm(100, mean = 1100, sd = 75))

qnorm(0.95, mean = new_mu, sd = 75)

```


## Q5 : Done
You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?

Bernouli's Distribution: **p^x.(1-p)^x**
Binomial Distribition = **n!/x!.(n-x)!**

```{r}
choose(5, 4)*0.5^4*(1-0.5)^1 + choose(5, 5)*0.5^5
```

## Q6

The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?

Mean: 15
SD: 10
p: ?
n: 100
Mean interval: 14-16


## Q7: Done

Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?

Mean: 0.5
Variance: 1/12
n = 1000

```{r}

sum(rnorm(1000, mean = 0.5, sd = sqrt(1/12)))/1000
```


## Q8: Done

The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?

Mean: 5
3 hrs watch
Probability of 10 ro few

```{r}
ppois(10, lambda = 5 *3)
```